---
title: "Unit1 Final Assignment Berries"
author: "Hao Shen"
date: "2020/10/17"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# 1 Summary

In this assignment, we do EAD for data set berries from USDA. In data cleaning process, we mainly separate mixed stings into several variables and delete variables with meaningless or replicated information. Then, according to data set attributions, we divided the whole date set into two parts:

* Bearing: a part contains information about farm chemical and fertilizer usage on berries.
* Market: a part with price, yield, area related information about berries.

For Bearing part, we use principal component analysis to explore variables' internal relationship and the results shows the PCA will not help a lot in regression.

For Market part, since the relationships between variables are extremely simple, we summarize 13 equations to explain their connections.

Besides, we deploy a shiny app [Berries-shiny](https://haoshen.shinyapps.io/Berries-shiny/) for data display and a slide [Berries-slide](Berries-slide.pdf) to presente the results.

# 2 Data cleaning

After initially rough observation of the berries data set downloaded from USDA [<sup>1</sup>](#refer-anchor-1) , we can find it generally consisted by two parts: 

* Bearing: one mainly about soil bearings on berries including farm chemical bearings like fungicide, herbicide, etc. and fertilizer bearings including nitrogen, potash, etc.
* Market: one contains market related data about berries including price, yield, acre, etc.

So, we believe the whole data sets should be separated into two part and analyzed respectively. Here we just use data of 'BLUEBERRIES' for example.

## 2.1 Select columns with useful information

First, we load data from local disk and filter those lines related to 'BLUBERRIES'. Then we drop some columns for the information they contained is either meaningless or replicated, and get a data frame of 7,597 observations with 7 variables summarized as follow:

```{R "Column Selection"}
berries=read.csv('Berries-data/berries.csv')
b_berry=filter(berries,Commodity=='BLUEBERRIES')
b_berry=select(b_berry,
               -Program,
               -(Period:Geo.Level),
               -(State.ANSI:Watershed),
               -CV....)
summary(b_berry)
```

## 2.2 Separate 'b_berry' into two tables

Now, we start to separate the whole data set into two tables bearing and market as mentioned above by detecting whether 'BEARING' occurred in Data.Item. 

Then in exploring remaining data or so called market related data, it is interesting to find 'BLUEBERRIES' from Maine State are labeled with WILD while others are all 'TAME'. This may caused by 'the wild blueberry, sometimes called the low-bush blueberry, is native to Maine and thrives in its glacier-churned soil and challenging seasons.' [<sup>2</sup>](#refer-anchor-2) However, for convenience, we will only label it in a new TAME column by 1 or 0 and ingore it in further analysis.

```{R "Separate 'b_berry' into two tables"}
#write.csv(b_berry,'b_berry.csv',row.names=F)
#unique(b_berry$Data.Item)
b_berry_bearing=b_berry[str_detect(b_berry$Data.Item,'BEARING'),]
b_berry_market=b_berry[!str_detect(b_berry$Data.Item,'BEARING'),]
b_berry_market$TAME[str_detect(b_berry_market$Data.Item,'TAME')]=1
b_berry_market$TAME[str_detect(b_berry_market$Data.Item,'WILD')]=0
```

## 2.3 Create table1 'b_berry_bearing'

Here we come into the most complicated part to separate Data.Item, for each of them includes at least 3 parts of information with only 1 string like 'BLUEBERRIES, BEARING - APPLICATIONS, MEASURED IN LB'

* Commodity='BLUEBERRIES': a part contains the same information of Commodity column which can be deleted directly.
* Measurement='BEARING - APPLICATIONS': a part denote the measurement used to generate Values. It can be either 'TREATED' or 'APPLICATION'.
* Unit='MEASURED IN LB': a part denote the unit of Values. For observations with 'APPLICATION', units include LB, LB/ACRE/APPLICATION, etc. For observations with 'TREATED', unit only include percentage.

Since both comma and space contained in strings and varies from observation to observation, it is hard to use normal separator link ',' and ' '. And we believe it can be solved by plugging our own separator '#' into the right places, and just split strings by these new separators.

```{R "Separate 'Data.Item' of 'Bearing'"}
#unique(b_berry_bearing$Data.Item)
b_berry_bearing$Data.Item=str_replace_all(b_berry_bearing$Data.Item,
                                          c('BLUEBERRIES, BEARING - '='',
                                            ', MEASURED IN '='#'))
b_berry_bearing=separate(b_berry_bearing,Data.Item,c('Measurement','Unit'),'#')
```

After separation of Data.Item, we notice that observations with 'CHEMICAL' Domain included are different from those with 'FERTILIZER' in both Domain and Domain.Category columns. Therefore in next steps we need to separate the whole bearing table into two sub-tables and handle their data respectively. 

```{R "Separate 'b_berry_bearing' into two tables"}
#unique(b_berry_bearing$Domain)
b_berry_chemical=b_berry_bearing[
  str_detect(b_berry_bearing$Domain,'CHEMICAL'),]
b_berry_fertilizer=b_berry_bearing[
  str_detect(b_berry_bearing$Domain,'FERTILIZER'),]
```

### 2.3.1 Create sub-table 'b_berry_chemical'

For sub-table with 'CHEMICAL' Domain, it has a Domain.Category like 'CHEMICAL, INSECTICIDE: (PHOSMET = 59201)', which contains two parts:

* Domain='CHEMICAL, INSECTICIDE': more specific information than Domain about chemical types.
* Category='(PHOSMET = 59201)'': a part with information of chemical components.

So, we decide to just delete original Domain column and separate the Domain.Category into two part, Domain and Category columns to gain contain more information with just 2 columns.

```{R "Separate 'Domain.Category' of 'b_berry_chemical'"}
b_berry_chemical=select(b_berry_chemical,-Domain)
#unique(b_berry_chemical$Domain.Category)
b_berry_chemical$Domain.Category=str_replace_all(b_berry_chemical$Domain.Category,
                                                 c('CHEMICAL, '='',
                                                   ': '='#'))
b_berry_chemical=separate(b_berry_chemical,Domain.Category,c('Domain','Category'),'#')
```

### 2.3.2 Create sub-table 'b_berry_fertilizer'

For sub-table with 'FERTILIZER' Domain, its Domain.Category is like 'FERTILIZER: (SULFUR)' which also has two parts:

* Domain='FERTILIZER': a part contains the same information as Domain column.
* Category='(SULFUR)'': a part with information of chemical components.

So, we can also delete original Domain column and separate the Domain.Category into two parts.

```{R "Separate 'Domain.Category' of 'b_berry_fertilizer'"}
b_berry_fertilizer=select(b_berry_fertilizer,-Domain)
#unique(b_berry_fertilizer$Domain.Category)
b_berry_fertilizer=separate(b_berry_fertilizer,Domain.Category,c('Domain','Category'),': ')
```

### 2.3.3 Form new table1 'b_berry_bearing'

Now, we can combine these two sub-tables to form the new bearing table. It is displayed and summarized as follow:

```{R "Form new 'b_berry_bearing'"}
b_berry_bearing=rbind(b_berry_chemical,b_berry_fertilizer)
head(b_berry_bearing)
summary(b_berry_bearing)
```

## 2.4Create table2 'b_berry_market'

Basically, the creation of table2 'b_berry_market' is the same as the creation of table1, with just a little changes in handling Data.Item column.

Here the most rows of Data.Item column have the same format of table1 'b_berry_chemical', consisted by Commodity, Measurement and Unit. But for some rows with Data.Item equal contains 'ACRES HARVESTED', it only have two parts of information, Commodity and Measurement. But it is early to handle, since if we separate them as others, the new Unit column will be filled by NA, which can be easily replaced by its real unit 'ACRE'.

After cleaning, it is displayed and summarized as follow:

```{R "Separate 'Data.Item' of 'b_berry_market'"}
b_berry_market=select(b_berry_market,-Domain,-Domain.Category)
#unique(b_berry_market$Data.Item)
b_berry_market$Data.Item=str_replace_all(b_berry_market$Data.Item,
                                       c('BLUEBERRIES, TAME'='#',
                                         'BLUEBERRIES, WILD'='#',
                                         ', MEASURED IN '='#'))
b_berry_market$Data.Item=str_replace_all(b_berry_market$Data.Item,
                                       c('# - '='',
                                         '#, '=''))
b_berry_market=separate(b_berry_market,Data.Item,c('Measurement','Unit'),'#')
#b_berry_market[is.na(b_berry_market$Unit),]
b_berry_market$Unit[is.na(b_berry_market$Unit)]='ACRE'
head(b_berry_market)
summary(b_berry_market)
```

## 2.5 Remove useless b_berries and save data

We finally get two data sets of bearings and market after all above processes and we need to save them at two place: one in Berries-data folder for further EDA and Slide making, one in Berries-shiny for Shiny app to deploy.

```{R "Remove useless b_berries"}
rm(b_berry_fertilizer,b_berry_chemical)
#for EDA & Slide
write.csv(b_berry_bearing,"Berries-data/b_berry_bearing.csv",row.names=F)
write.csv(b_berry_market,"Berries-data/b_berry_market.csv",row.names=F)
#for Shiny
write.csv(b_berry_bearing,"Berries-shiny/b_berry_bearing.csv",row.names=F)
write.csv(b_berry_market,"Berries-shiny/b_berry_market.csv",row.names=F)

```

# 3 EDA for 'bearing'

EDA for 'bearing', especially for the chemical parts of bearing is extremely hard for a normal person since we do not have so much understanding of various farm chemicals. However, we can still analyze parts of them like total FUNGICIDE, HERBICIDE, INSECTICIDE, Others and total NITROGEN, PHOSPHATE, POTASH, SULFUR. Here we just use Principal Component Analysis (PCA) to analyze their relationships. Since the FERRTILIZER data only collected in 2015 and 2019, we will choose these two years only. Besides, for missing value, we use mean value to replace.

```{r}
# load data
bearing=read.csv('Berries-data/b_berry_bearing.csv')
bearing_value=filter(bearing,
                     Year%in%c('2019','2015'),
                     Unit=='LB',
                     Category%in%c('(NITROGEN)','(PHOSPHATE)','(POTASH)',
                                   '(SULFUR)','(TOTAL)')
                     )
# arrange data
bearing_value$DC=paste(bearing_value$Domain,bearing_value$Category)
bearing_value=select(bearing_value,DC,Value)
bearing_value=arrange(bearing_value,bearing_value[,1])
# handle missing value
for(i in unique(bearing_value$DC)){
  m=bearing_value$Value[bearing_value$Value!=' (D)'
                        &bearing_value$DC==i]%>%
    str_replace_all(c(','=''))%>%
    as.numeric()%>%
    mean()
  bearing_value$Value[bearing_value$Value==' (D)'
                    &bearing_value$DC==i]=m

}
# transform to numeric
bearing_value$Value=as.numeric(str_replace_all(bearing_value$Value,c(','='')))
# transform for PCA
j=1
for(i in unique(bearing_value$DC)){
  if(j==1)bearing_pca=bearing_value$Value[bearing_value$DC==i]
  else bearing_pca=cbind(bearing_pca,bearing_value$Value[bearing_value$DC==i])
  j=0
}
colnames(bearing_pca)=unique(bearing_value$DC)
head(bearing_pca)
```

After getting the data, we use prcomp function for PCA and summary its components, cumulative proportion and plot [<sup>3</sup>](#refer-anchor-3) as follow:

```{r}
pca=prcomp(bearing_pca,center = T,scale. = T)
summary(pca)
print(pca)
```
```{r}
devtools::install_github("vqv/ggbiplot")
ggbiplot::ggscreeplot(pca) 
```

As we can see, the first 4 principal component is important. Considering there is only 8 initial variables, PCA is not so help in regression.

# 4 EDA for 'market'

The EDA for 'market' table is much easier. The relationships between different variables are extremly clear. We can use an example of California in 2019 to illustrate these:

```{R}
market=read.csv('Berries-data/b_berry_market.csv')
market_value=filter(market,
                    Year=='2019',
                    State=='CALIFORNIA')
market_value=select(market_value,-Commodity,-TAME)
market_value=arrange(market_value,market_value[,3])
(market_value=data.frame(No.=1:nrow(market_value),market_value))
write.csv(market_value,'1.csv')
```

So, here there are 13 equations to explain there relationships:

*   Row1    =   Row10   /   Row13
*   Row2    =   Row3    /   Row4
*   Row3    =   Row2    *   Row4
*   Row4    =   Row3    /   Row2
*   Row5    =   Row10   -   Row12
*   Row6    =   Row11   /   Row12
*   Row7    =   Row8    /   Row9
*   Row8    =   Row7    *   Row9
*   Row9    =   Row8    /   Row7
*   Row10   =   Row5    +   Row12
*   Row11   =   Row6    *   Row12
*   Row12   =   Row4    +   Row9
*   Row13   =   Row10   /   Row1

# 5 Create Shiny

Then, we create a shiny app to display the data we get. The app [Berries-shiny](https://haoshen.shinyapps.io/Berries-shiny/) has been deployed on shinyapps.io. And the code is listed in [app.R](Berries-shiny/app.R).

# 6 Create Slide
Finally we create a slide [Berries-slide](Berries-slide.pdf) to summary our whole EDA process. The code of slide is in [Berries-slide.Rmd](Berries-slide.Rmd).

# 7 Reference
<div id="refer-anchor-1"></div>
[1] [National Agricultural Statistics Service](https://quickstats.nass.usda.gov/)
<div id="refer-anchor-2"></div>
[2] [Visit Maine](https://visitmaine.com/things-to-do/agricultural-attractions/wild-for-blueberries)
<div id="refer-anchor-3"></div>
[3] [Vince Vu](https://github.com/vqv/ggbiplot)
