---
title: "Unit1 Final Assignment Berries"
author: "Hao Shen"
date: "2020/10/17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# 1 Summary

# 2 Data cleaning

After initially rough observation of the berries data set downloaded from USDA, we can find it generally consisted by two parts: 

* one mainly about soil bearing on berries including farm chemical bearings like fungicide, herbicide, etc. and fertilizer bearings including nitrogeon, potash, etc.
* 

## 2.1 Select columns with useful information
```{R "Column Selection"}
berries=read.csv('Berries-data/berries.csv')
b_berry=filter(berries,Commodity=='BLUEBERRIES')
b_berry=select(b_berry,
               -Program,
               -(Period:Geo.Level),
               -(State.ANSI:Watershed),
               -CV....)
```

## 2.2 Separate 'b_berry' into two tables
```{R "Separate 'b_berry' into two tables"}
#write.csv(b_berry,'b_berry.csv',row.names=F)
#unique(b_berry$Data.Item)
b_berry_bearing=b_berry[str_detect(b_berry$Data.Item,'BEARING'),]
b_berry_market=b_berry[str_detect(b_berry$Data.Item,'TAME')|
                         str_detect(b_berry$Data.Item,'WILD'),]
b_berry_market$TAME[str_detect(b_berry_market$Data.Item,'TAME')]=1
b_berry_market$TAME[str_detect(b_berry_market$Data.Item,'WILD')]=0
```

## 2.3 Create sub-table 'b_berry_bearing'
```{R "Separate 'Data.Item' of 'Bearing'"}
#unique(b_berry_bearing$Data.Item)
b_berry_bearing$Data.Item=str_replace_all(b_berry_bearing$Data.Item,
                                          c('BLUEBERRIES, BEARING - '='',
                                            ', MEASURED IN '='#'))
b_berry_bearing=separate(b_berry_bearing,Data.Item,c('Measurement','Unit'),'#')
```

```{R "Separate 'b_berry_bearing' into two tables"}
#unique(b_berry_bearing$Domain)
b_berry_chemical=b_berry_bearing[
  str_detect(b_berry_bearing$Domain,'CHEMICAL'),]
b_berry_fertilizer=b_berry_bearing[
  str_detect(b_berry_bearing$Domain,'FERTILIZER'),]
```

### 2.3.1 Create sub-table 'b_berry_chemical'
```{R "Separate 'Domain.Category' of 'b_berry_chemical'"}
b_berry_chemical=select(b_berry_chemical,-Domain)
#unique(b_berry_chemical$Domain.Category)
b_berry_chemical$Domain.Category=str_replace_all(b_berry_chemical$Domain.Category,
                                                 c('CHEMICAL, '='',
                                                   ': '='#'))
b_berry_chemical=separate(b_berry_chemical,Domain.Category,c('Domain','Category'),'#')
```

### 2.3.2 Create sub-table 'b_berry_fertilizer'
```{R "Separate 'Domain.Category' of 'b_berry_fertilizer'"}
b_berry_fertilizer=select(b_berry_fertilizer,-Domain)
#unique(b_berry_fertilizer$Domain.Category)
b_berry_fertilizer=separate(b_berry_fertilizer,Domain.Category,c('Domain','Category'),': ')
```

### 2.3.3 Form new 'b_berry_bearing'
```{R "Form new 'b_berry_bearing'"}
b_berry_bearing=rbind(b_berry_chemical,b_berry_fertilizer)
```

## 2.4Create sub-table 'b_berry_market'
```{R "Separate 'Data.Item' of 'b_berry_market'"}
b_berry_market=select(b_berry_market,-Domain,-Domain.Category)
#unique(b_berry_market$Data.Item)
b_berry_market$Data.Item=str_replace_all(b_berry_market$Data.Item,
                                       c('BLUEBERRIES, TAME'='#',
                                         'BLUEBERRIES, WILD'='#',
                                         ', MEASURED IN '='#'))
b_berry_market$Data.Item=str_replace_all(b_berry_market$Data.Item,
                                       c('# - '='',
                                         '#, '=''))
b_berry_market=separate(b_berry_market,Data.Item,c('Measurement','Unit'),'#')
#b_berry_market[is.na(b_berry_market$Unit),]
b_berry_market$Unit[is.na(b_berry_market$Unit)]='ACRE'
```

## 2.5 Remove useless b_berries and save data
```{R "Remove useless b_berries"}
rm(b_berry_fertilizer,b_berry_chemical)
#for EDA & Slide
write.csv(b_berry_bearing,"Berries-data/b_berry_bearing.csv",row.names=F)
write.csv(b_berry_market,"Berries-data/b_berry_market.csv",row.names=F)
#for Shiny
write.csv(b_berry_bearing,"Berries-shiny/b_berry_bearing.csv",row.names=F)
write.csv(b_berry_market,"Berries-shiny/b_berry_market.csv",row.names=F)

```

# 3 EDA for 'bearing'
```{r}
bearing=read.csv('Berries-data/b_berry_bearing.csv')
head(bearing)
bearing_value=filter(bearing,Unit=='LB')
unique(bearing_value$Domain)

name_category=NA
j=1
for (i in unique(bearing_value$Category)) {
  nrow_category=nrow(filter(bearing_value,Category==i)
                        )
  if (nrow_category>=16){
    name_category[j]=i
    j=j+1
  }
}

miss_data=NA
x=filter(bearing_value,Category%in%name_category)

```

```{r}
(b_domain=unique(bearing$Domain))
unique(filter(bearing,Domain==b_domain[1])$Year)
unique(filter(bearing,Domain==b_domain[2])$Year)
unique(filter(bearing,Domain==b_domain[3])$Year)
unique(filter(bearing,Domain==b_domain[4])$Year)
unique(filter(bearing,Domain==b_domain[5])$Year)
```
# 4 EDA for 'market'
```{R}
market=read.csv('Berries-data/b_berry_market.csv')
market=filter(market,Measurement!='NOT HARVESTED - PRODUCTION')
index_measure=data.frame(No=1:nrow(unique(select(market,Measurement,Unit))),
                         unique(select(market,Measurement,Unit)))
```

# 5 Create Shiny
# 6 Create Slide
